\documentclass{howto}
\usepackage{moreverb}                   %% Verbatim Code Listings

\title{Cheetah Beginner's Guide}
\release{0.9.12 post b2}

\author{The Cheetah Development Team}
\authoraddress{\email{cheetahtemplate-discuss@lists.sourceforge.net}}

\begin{document}
\maketitle


\tableofcontents

\copyright{Copyright 2002, The Cheetah Development Team.
      This document may be copied and modified under the terms of the
      {\bf Open Publication License} \url{http://www.opencontent.org/openpub/} }

This document is not written yet.  It is intended to be an introduction to
Cheetah for template maintainers who don't know Python.  The Users' Guide is
the definitive specification for Cheetah and explains how to install and use
it, but it assumes a knowledge of Python and describes Cheetah in terms of
Python.  

For now, just try your best to follow the Users' Guide, and ask questions if
you don't get something.  {\bf Let us know of things you wish were in the
Beginners' Guide, so we can keep them in mind when we write it.}

If you would like to volunteer to write this Beginners' Guide, let us know.
It needs a writer who knows Python but trains Cheetah users who don't, so
(s)he knows what their concerns are.

    %% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %\include{introduction}
    %\include{gettingStarted}
    %\include{howItWorks}

    %% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    %\appendix

\end{document}

% Local Variables:
% TeX-master: "users_guide"
% End:      
