\documentclass{howto}
\usepackage{moreverb}                   %% Verbatim Code Listings

\title{Cheetah Developers' Guide}
\release{0.9.13b1}

\author{Mike Orr}
\authoraddress{\email{iron@mso.oz.net}}

\begin{document}
\maketitle


%\tableofcontents

\copyright{Copyright 2002, Mike Orr.
      This document may be copied and modified under the terms of the
      {\bf Open Publication License} \url{http://www.opencontent.org/openpub/} }

This document is still being written.  Material to include:

** Show how cheetah-compile translates caching, filter, the other directives,
and comments.

** List all the methods accessible via 'self'.  (Some of this may belong in the
Users' Guide.)

** Describe how to commit changes to CVS (assuming you have a login/password),
and how to run the test suite beforehand.

** Describe how Cheetah is packaged in distutils format, and how the regression
test suite works.

** History of Cheetah (including the WebOnions image).  Describe major
milestone changes from the changelog.

% @@MO: WebOnions logo.  (Goes with history.)

** Design decisions and tradeoffs that have been made.

** How to build the documentation.  Why LaTeX, a minimum LaTeX reference, etc.

** Show where you can omit the '\$' (leniency) and where you can't.

** Appendix A will be a formal BNF grammar of Cheetah.

% Two files placeholders.tex and safeDelegation.tex contain material to be
%incorporated into this guide.


    %% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \include{introduction}
    \include{files}
    \include{template}
    \include{compiler}
    \include{pyModules}
    \include{placeholders}
    \include{comments}
    \include{output}
    \include{inheritanceEtc}
    \include{flowControl}
    \include{errorHandling}
    \include{parserInstructions}

    \include{parser}
    \include{compiler2}
    \include{design}
    \include{patching}
    \include{documenting}
    
    %% %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \appendix
    \include{bnf}
    \include{safeDelegation}
\end{document}

% Local Variables:
% TeX-master: "users_guide"
% End:      
