\section{Developers Information}
\label{developer}

This chapter is for people on the Cheetah development team, and for those who
want the ``inside scoop'' on how Cheetah works.

See the TODO file in the Cheetah source distribution for features we're
working on or would like to add.  In the CVS version of Cheetah, the directory
docs/devel contains text files detailing possible implementation strategies.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Adding features to Cheetah}

Whenever you add a feature to Cheetah, follow these steps:

\begin{enumerate}
     
\item Discuss its need and design on the email list and get approval.
     
\item Add the feature.
     
\item Write regression tests for it.  Verify the test suite does not break.
     
\item Add an entry to the CHANGES file explaining what you did.
     
\item Remove the entry on the TODO list if present.
     
\item Update the documentation.
     
\item Commit it all.
     
\item Notify the list and ask for people to test it.

\end{enumerate}


%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Running regression tests}

Cheetah comes with a set of regression tests to verify it's working properly.
To run the tests:

\begin{enumerate}
     
\item \code{cd} into \code{src/Tests} in the Cheetah source distribution.
     
\item Run the command: \code{python Test.py}
     
\item It will run the test cases and then print either "OK" or an error message.

\end{enumerate}

The test suite tries to write the file \code{parseTest.txt} in the current
directory if it doesn't exist.  You will get an IOError if you don't have write
permission.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Adding regression tests}

The test cases are divided into several modules:
\code{Cheetah.Tests.SyntaxAndOutput}, \code{Cheetah.Tests.NameMappper}, and
\code{Cheetah.Tests.Template}. If you have discovered a bug please make a test
case following the format in those modules and submit it for inclusion.

% Local Variables:
% TeX-master: "users_guide"
% End:      
